Application-Specific Morphological Analyzer for Arabic Text Mining

### 2011-04-30 trunk/tex/tp.tex
wrote automatic code for splitting prefixes, not useful yet for suffixes.
Did much of the work and splitted prefixes except for some documented issues due to subject/object modifications and discovered some minor inconsistencies in Buckwalter's work

@@

\begin{abstract}
Text Mining 
%can be considered as a subfield of both Natural Language Processing (NLP) and Data Mining.
is interested in automatically extracting useful information and relations from text documents.
%Extracted information is hard to discover manually or previously unknown
%Natural Language Processing applications perform 
%to discover useful information that is otherwise unknown.
The complexity of natural languages introduces interesting challenges 
to automatic analysis of underlying textual data.
Morphological analysis is key to many Natural Language Processing Applications.
General morphological analyzers such as Buckwalter and ElixirFM 
generate all morphological solutions.
However, this exhaustive enumeration may not be needed or appropriate for the
application at hand.
We propose developing Sarf, an Arabic application-specific morphological analyzer whose functionality can be customized 
based on the specific needs of the application. 
%This flexibility has been motivated by 
%the broad spectrum of text mining applications, each with its specific characteristics and corpora, which
%necessitate a special treatment.
Sarf encodes its lexicon into a non-deterministic composition of three finite automata.
It also solves the issue of `run-on' words, supports disambiguation using diacritics, 
handles multi-word expressions and performs online tokenization.
In our thesis, we will use Sarf to solve two useful Natural Language Processing applications: 
(1) the extraction of time entities from Arabic documents, and 
(2) the automation of the process of hadith authentication.

@@

The world we live in overwhelms us with electronic data. There is an ever-increasing gap between the amount
of information available and our ability to understand and analyze it. ``Omnipresent computers make
it too easy to save things that previously we would have trashed. Inexpensive disks
and online storage make it too easy to postpone decisions about what to do with all
this stuff -- we simply get more memory and keep it all.''~\cite{Witten:11}
This explosion in available data has motivated the automation of the task 
of extracting information of interest for a certain usage. 
Manual extraction is practically unfeasible due to
our limited precious time recourses compared to the magnitude of the task.
In addition, it was shown that automatic methods have been able to discover previously unknown 
information through relating separate datasets~\cite{JNi06}.

@@

Clearly, a small percentage of the data available worldwide 
is stored in databases in structured form; instead, the greater amount of information is available in the 
form of electronic text files and documents such as news articles, publications, governmental records, etc. ...
Even old paper documents are being ported to digital form to facilitate their use. 
Text Mining emerged as an attempt to automate the extraction of interesting, 
previously unknown and potentially useful information
from the huge amounts of text documents~\cite{Witten:08}.
Information Extraction (IE) from natural languages, however, is challenged
by the difficulty of getting structured 
entities from the underlying unstructured textual data.

%Information Extraction (IE) from Natural Language Texts has always been an important 
%problem in Data mining. What distinguishes text mining from the 
%other subfields of data mining is mainly due to the challenge of getting structured 
%entities from the underlying unstructured textual data. Clearly, a small percentage of the data available worldwide 
%is stored in databases in structured form; instead the greater amount of information is available in the 
%form of electronic text files and documents such as news articles, publications, governmental records, etc. ...
%Even old paper documents are being ported to digital form to facilitate their use.

Natural Language Processing (NLP)
as a field of research is interested in achieving a computational model of the language
and understanding its underlying meaning~\cite{sharp:01}. It is known that this task
is very complex and possibly not completely achievable. However, in most applications it 
is not needed to grasp the whole meaning of the text but to extract some useful entities.
This Information Extraction(IE) problem can be performed through some statistical approaches 
and/or partial NLP techniques.

A subproblem of NLP is morphology.
Morphological analyzers%~\cite{Sughaiyer:04}
are interested in understanding the internal structure of a word;
It considers words as being composed of 
several {\em morphemes}. 
A morpheme is a {\em stem} or an {\em affix}.
An affix is a {\em prefix, suffix,} or an {\em infix}.
Prefixes and suffixes are attached to the beginning and end of the word respectively; 
while infixes introduce changes in the stem itself. 
Morphological analyzers have been extensively used as building blocks 
in many text mining applications~\cite{Sou07}.  In Arabic, many morphological analyzers exist including:  
Buckwalter~\cite{Buckwalter:02},
Beesley~\cite{Beesley:01},
SAMA ~\cite{Kulick:10},
ElixirFM~\cite{Otakar:07}, 
AMIRA~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, 
and MADA+TOKAN~\cite{Habash:09}.
%Arabic however has distinctive characteristics that
%complicate its analysis; and consequently, much remains to be performed in the field.

In this thesis, we propose to develop Sarf, an application-specific 
morphological analyzer for the Arabic language 
which is capable of benefiting
from the specific characteristics of the application at hand. 
In addition, it supports disambiguation using diacritics and 
handles `run-on' words. We also propose to use Sarf to solve two interesting .
text mining applications: time entity extraction and hadith authentication.

\section{Applications of Text Mining}
\label{sec:Applications}

Research in text mining started in the mid 1980s when Swanson 
realized that slicing and combining seemingly unrelated medical 
articles led to the discovery of new 
hypotheses~\cite{JNi06}. 
Text Mining has gained much interest lately due to its wide range of applications,
including its use in the areas of advertisement, 
political campaigning, and other businesses.
However, these applications have substantially differing corpora (terminology, 
sentence structure, etc...). In this section, we present some of the 
previously explored applications of text mining.

\subsection{Sentiment Analysis} 
	This field -- also known as Opinion Mining' -- attempts classifying the polarity of sentiments of people 
	or even their emotional state ~\cite{Pang:08}. The inputs for such analyzers include product and movie reviews, 
	Internet blogs, electronic social networks and possibly newspapers. Early research was conducted on review sites
	to automate the task classifying whether the review was positive and negative. Consequently, users can get concise helpful
	sentiment summaries about the product, movie, etc...~\cite{Pang:02, Turney:02}. Such abilities motivated research targeting 
	the following applications:
	\subsubsection{Marketing Applications} 
		The strength of blogs in disseminating opinions is a two-edged sword that can 
		``make or break a company's reputation''~\cite{hoffman:08a}. Consequently, analyzing customers' opinions as portrayed 
		on online discussion forums should be detected as soon as possible. This input is vital in the company's policy for its
		reputation management. It can accordingly take actions to retain current customers, attract new ones and win back former 
		customers~\cite{hoffman:08b}. Sentiment Mining has also been used to predict customer churns~\cite{marketing:07,marketing:08}
		to take appropriate actions. It can help also in producing more effective advertisements based on the interests of the market.
	\subsubsection{Political Applications} 
		Political experts and analysts can benefit also. Since opinions extracted from textual data 
		can reveal much even without the need of performing any explicit surveys. According to ~\cite{Gryc:10}, it ``opens up many avenues
		for more accurate and naturalistic large-scale political analysis.'' Some research has been performed on classifying and predicting
		political inclinations~\cite{Durant:07, Yu:08}. This field can help political leaders to better understand the views of their people 
		and possibly perform better campaigns.

\subsection{Academic Applications} 
	In scientific disciplines where written papers and publications contain technical information, the need to retrieve
	specific information that a user needs, aids in \textbf{\em developing advanced search engines}. 
	Text mining plays a great role toward this end through the following:
	\subsubsection{Text Summarization} 
		When an application is interested in mining a text for the most important and informative portions,
		the application is considered to be tackling the field of `Text Summarization'. Summaries may have different forms such as being 
		a group of representative words, to a coherent abstraction of the text in novel terms and phrases that capture its main meaning. 
		The main motivation
		for such a research is to enhance search engines. By capturing the most informative concepts in a document, it aids determining a 
		better metric for a document's relevance to a search query. It also helps search engines display a shortened preview of the document for returned
		results. This reduces the overwhelming amount of information a user needs to read before deciding weather the whole document deserves
		comprehensive study. It is also helpful when the editorial staff of a magazine or newspaper need to present an article in the limited space.
		Similarly, summarization is vital when the content of 
		news articles needs to be dramatically shortened to be sent in an SMS form to subscribed PDA users. 
		Useful surveys of current research on text Summarization can be
		found in ~\cite{Radev:02,Das:07}.
	\subsubsection{Text Categorization and Clustering} 
		In many cases, it is very useful to be able to search for documents by theme, topic, domain, etc...
		This ability is achieved if documents can be automatically clustered into different categories over which a user can search. It has been also 
		proposed to cluster documents based on biographical data~\cite{Kostoff:99}. Technically speaking the difference between Categorization and
		Clustering is that the former organizes documents into predefined categories where in the latter the categories arise from the document content
		themselves~\cite{sharp:01}. Many approaches exist in this field as described in~\cite{Berry:03, Aas:99, MUNTEANU:05}.
		\begin{enumerate}
		\item \textbf{Email Classification:} It is noteworthy to mention that it is text categorization can be used for emails 
			-- such as urgent, work, friends, spam, etc... Consequently, it can be used for spam detection and email 
			filtering~\cite{Mock:01,bekkerman:04}.
		\item \textbf{Parental Control:} It has also been proposed to use categorization for parental control by
			automatically detecting inappropriate content for their children~\cite{Kontostathis:09}.
		\end{enumerate}
	\subsubsection{Entity Extraction:} Named Entity Extraction/Recognition (NER) targets being able to detect phrases that convey some general concept
		and/or belong to a predefined set of categories such as persons, places, time, etc... It has been argued that having the luxury of searching
		for entities enhances search abilities instead of being restricted to a list of words appearing in a text~\cite{basis:06}. 
		It seems that humans think of their
		queries in terms of entities/concepts and then try translating them to some list of words which are most likely to appear in a document that
		is relevant to their search query. In general, a user does not ``know all the search terms that will return all instances of the desired
		information''~\cite{basis:06}. A survey of research in the field is presented in ~\cite{Nadeau:07}.

\subsection{Security Applications} 
	It has been proposed that text mining of emails can be used for detecting money laundering and 
	terrorist plans. However the effectiveness of data mining in this context is still debatable~\cite{Jonas:06,Federici:07}.

	We also propose another security application which can be used to be used when investigating in criminal cases:
	Given two sets of reports consisting of crime scene investigation and suspect interrogation reports, 
	an officer may query use a specialized text mining application to query for relations between suspects in terms of 
	criminal action, locality, or third party persons unknown to the officer.
	It can be envisioned that such an application will return two convicted people who had no direct relation
	in the reports, but who are related only through a third person unknown to the investigator. 
	Such a query is very hard to answer manually.

\subsection{Medical Applications} 
	Medical records and clinical reports are studied to detect
	patterns of occurrence of certain symptoms and diseases. It attempts relate those patterns with
	other biological factors such as certain 
	genes or protein interaction~\cite{blaschke:99, Krallinger:05, Rebholz-Schuhmann:07, Cohen:08, Tsuruoka:08}.
	They are also used to analyze factors correlating with medical accidents to minimize their occurrence~\cite{Kimura:08}.
	Till present, most of the interest in text mining comes from biological sciences~\cite{Giannopoulou:08}.

\subsection{Hadith Authentication}

In this thesis, we will present and propose to solve an interesting application 
for text mining in the field of Islamic Studies. 
A \noVocRL{.hady_t}~\footnote{In this document, 
we use the default ArabTeX transliteration style ZDMG.}
is a narration related to the prophet Mohammad
through a \noVocRL{sanad} or a sequence of narrators. 
The collections of traditions are the second source of
jurisprudence after the \noVocRL{qor'An} for all Islamic schools of thought. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in 
Arabic with its transliteration and translation. 
We show proper names in boxes connected
to form complex names of narrators. 
For example, 
\novocalize
\noTrRL{qtybT} is the first name 
of narrator $n_1$, and 
\noTrRL{s`yd} is the name of his father as 
the word \noTrRL{bn} (son of) indicates. 
The sequence of names from $n_1$ to $n_5$ 
constitutes the \noArRL{sanad} 
of the \noArRL{.hady_t}. 
The second part of the \noArRL{.hady_t} is the 
\RL{matn} (content) and constitutes the actual content 
of the tradition.

@@

@@

	Morphological ambiguity is a `notorious' problem for Arabic language~\cite{Kiraz:98}.
	Ambiguity refers to the fact that one word may lead to several possible solutions.
	As discussed in~\cite{Attia:08a}, the following are the reasons of genuine ambiguity\footnote{
	Here we are discussing ambiguity that is inherent to the language. Other sources of ambiguity
	may be due to over-approximations and relaxed compatibility rules in the morphological analyzer.} 
	in the language:
	\begin{enumerate}
		\item \textbf{Orthographic alternation operations:} These refer to alterations from the stem
			original letters upon inflection, which can make it similar to other distinct words.
			For instance, \noVocRL{ya`id} can result in 5 different solutions, 2 of which are
			inflections of \RL{A`id} (bring back) and \RL{`od} (come back) as in \RL{lam ya`od}.
			~\cite{Attia:08a}
		\item \textbf{Prefixes and Suffixes that are homomorphic:} For example, 
			``the prefix {\em ta-} can indicate 3rd person feminine or 2nd person masculine''~\cite{Attia:08a}.
			Consider \noVocRL{tadros} which can mean `she studies' or `you (masculine) study'.
		\item \textbf{``Coincidental Identity'':} In this case, complex forms (i.e. containing affixes) coincidentally
			are homomorphic to a simple full word ~\cite{Kamir:02}. 
			%An example is in the homomorphism of \noVocRL{Akala}
			%he ate) and \noVocRL{Akalla} [\noVocRL{A} + \noVocRL{kalla}] (he made tired).
			\vocalize 
			For instance, the word \RL{'a.hmadH} may have two valid morphological analyses. 
			The word means ``I praise him'' when the letter \RL{'a} is a prefix and  it means
			``his Ahmad'' when the letter \RL{'a} is part of the proper noun stem \RL{'a.hmad}.
		\item \textbf{Natural Homomorphism in Un-inflected Forms:} This occurs in \noVocRL{_dahaba} which can mean
			`went' as a verb and `gold' as a noun.~\cite{Attia:08a}
		\item \textbf{Omitted Diacritizations} (will be discussed next)
	\end{enumerate}

	The statistical measure of ambiguity is the average count of valid solutions per word. According to our results performed
	on 3 books of hadith~\cite{IbnHanbal,AlKulayni,AlTousi}, the resulting ambiguity score was 2.62 using an extended corpus to that of
	Buckwalter Analyzer~\cite{Buckwalter:02}\footnote{This analyzer will be considered in Section \ref{sec:literature}.}.
	Those results where verified by ~\cite{Attia:06b} which reported a similar score (2.6) and a score of 1.76 for English.
	This difference can be used to argue that Arabic is more ambiguous than English. Still however, the author of ~\cite{Attia:06b}, Attia,
	claims that this ambiguity resulted from some grammatical over-approximations and inaccuracies in the lexicon of Buckwalter along with
	its inclusion of classical terms that are rarely used. By this he means that the score does not reflect the intrinsic Arabic ambiguity and 
	they report a score of 1.75 for their tool which used tighter grammar and lexical rules and omitted classical words. Although, Dr. Attia
	has a point relating to exaggerated numbers due to over-approximations, the reduction of the lexicon is not a correct approach since it does
	not reflect the ambiguity of Arabic but of the subset of Arabic needed for some applications (mainly modern text documents such as news).
	In all cases, this is a question open to future investigations by NLP researchers and Arabic linguists...

@@

	Since most Arabic documents are unvocalized, the ambiguity of the morphological analyses is significantly increased.
	For instance, in many cases, it is impossible to distinguish between active and passive verbs on the word level without
	appropriate diacritics such as the undiacriticized versions of \RL{`abada} ($<$subject$>$ worshiped) and \RL{`obida} ($<$subject$>$ was worshiped).

@@

	Arabic letters can have up to 
	four different forms
	corresponding to their position in a word, i.e. beginning,
	middle, end of word and separate forms. 
	This allows the phrase
	\noTrnoVocRL{il_A\nospace almdrsT}  with no delimiter within
	to be visually recognizable
	as two separate words \noVocRL{il_A} (to) and \noVocRL{almdrsT} (the school).
	This happens because the first word \noVocRL{il_A} ends with
	\noVocRL{_A} a non-connecting letter. 
	The non-connecting letters are \RL{A}, \noVocRL{'a}, \noVocRL{'i},
		\noVocRL{'A},\noVocRL{_A}, \noVocRL{w}, \noVocRL{|u'|}, \noVocRL{|i'|},	
		\noVocRL{d}, \noVocRL{_d}, \noVocRL{r}, \noVocRL{z} and \noVocRL{T}.
	These words occur often in text and greatly increase the
	difficulty of tokenization and are referred to as 
	``run-on'' words~\cite{Buckwalter:04}.
	Such difficulties are inherent in the 
	morphological analysis of the Arabic language and
	render the solutions inaccurate.
	Consequently, tokenization of Arabic text is not a trivial task of 
	detecting delimiters as is the case in other languages.

@@

The rest of the proposal is organized as follows. In Section 2, we present the problem(s) to be attempted. 
In Section 3, we review related literature. 
Then, in Section 4, we describe our preliminary results. 
Finally, in Section 5, a time-schedule and a research plan are presented.

@@

In other words, if the analyzer can be controlled at every step by an application-specific
controller, the analysis can be enhanced and reduced significantly.

>>>>

\renewcommand{\labelenumi}{\arabic{enumi}.}
\renewcommand{\labelenumii}{\labelenumi\arabic{enumii}.}
\renewcommand{\labelenumiii}{\labelenumii\arabic{enumiii}.}

@@

\begin{abstract}
Text Mining 
is interested in automatically extracting useful information and relations from text documents.
This information is hard to discover manually or is previously unknown.
This thesis is concerned with Arabic Text Mining where
%The complexity of natural languages introduces interesting challenges 
%to automatic analysis of underlying textual data.
morphological analysis is key.%to many Natural Language Processing Applications.
General-purpose morphological analyzers such as Buckwalter and ElixirFM exist for the Arabic language and
generate all possible morphological solutions.
Many Natural Language Processing applications need only few of these solutions.
We propose building Sarf, an Arabic application-specific morphological analyzer. 
Natural Language Processing applications can easily customize Sarf 
to prune unwanted solutions. This refinement allows applications to run more efficiently. 
%The flexible API of Sarf enables developers without expertise in morphology to have access rich NLP tools. (TODO: re-write)
Sarf encodes its lexicon into a non-deterministic composition of three finite transducers.
It also solves the issue of `run-on' words, supports disambiguation using diacritics, 
handles multi-word expressions and performs online tokenization.
We will apply Sarf to  
(1) extract time entities from Arabic documents,
(2) extract narrators from hadith documents,
(3) relate narrators to their biographies in separate documents, and 
 thus automate the process of hadith authentication.

@@

The world we live in overwhelms us with electronic data. There is an ever-increasing gap between the amount
of information available and our ability to understand and analyze it. ``Omnipresent computers make
it too easy to save things that previously we would have trashed. Inexpensive disks
and online storage make it too easy to postpone decisions about what to do with all
this stuff -- we simply get more memory and keep it all.''~\cite{Witten:11}
Even old paper documents are being ported to digital form to facilitate their use. 
This explosion in available data has motivated the automation of the task 
of extracting information of interest for a certain usage. 
Manual extraction is practically unfeasible due to
our limited precious time recourses compared to the magnitude of the task.
In addition, it was shown that automatic methods have been able to discover previously unknown 
information through relating separate datasets~\cite{JNi06}.

@@

While a small percentage of data %available worldwide 
is available in structured form, the greater amount is stored in unstructured form such as  
text files.
Examples are news articles, publications, governmental records, etc. ...
{\em Text Mining} emerged as an attempt to automate the extraction of interesting, 
previously unknown and potentially useful information
from the huge amounts of text documents~\cite{Witten:08}.
%Information Extraction (IE) from natural languages, however, is challenged
%by the difficulty of getting structured 
%entities from the underlying unstructured textual data.
Natural languages are inherently ambiguous which renders the problem of 
Information Extraction (IE) from natural language texts harder than 
its counterpart in structured data.

Natural Language Processing (NLP), one of the hardest fields in Artificial Intelligence,
is interested in building a computational model of text or speech
and their underlying meaning~\cite{sharp:01}. 
%It is known that this task
%is very complex and possibly not completely achievable. 
Many applications target entities and relations that can be 
easily extracted with partial computational models derived from
statistical and morphological analysis techniques.
%needed to grasp the whole meaning of the text but to extract some useful entities.
%This IE problem can be performed through some statistical approaches 
%and/or partial NLP techniques.

%A subproblem of NLP is morphology.
Morphological analyzers %~\cite{Sughaiyer:04}
consider the internal structure of a word and compose it from
several {\em morphemes}. 
A morpheme is a {\em stem} or an {\em affix}.
An affix is a {\em prefix, suffix,} or an {\em infix}.
Prefixes and suffixes are attached to the beginning and end of the word respectively; 
while infixes introduce changes in the stem itself. 
Morphological analyzers have been extensively used as building blocks 
in many text mining applications~\cite{Sou07}.  In Arabic, morphological analyzers such as  
Buckwalter~\cite{Buckwalter:02},
Beesley~\cite{Beesley:01},
SAMA ~\cite{Kulick:10},
ElixirFM~\cite{Otakar:07}, 
AMIRA~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, 
and MADA+TOKAN~\cite{Habash:09}
exist.
They preprocess text to remove diacritics, short vowels, such as 
\RL{ha}, \RL{hu}, \RL{hi}, \RL{TaN}, \RL{huN}, \RL{hiN}.~\footnote{In this document, 
we use the default ArabTeX transliteration style ZDMG.} Diacritics help to disambiguate the
meaning of a word. For example, \noVocRL{wrd} may mean ``rose"(\RL{ward}), ``came"(\fullvocalize\RL{warada}) 
or ``and he replied"(\RL{waradda}). We think that using diacritics in the analysis can highly increase accuracy.
Current analyzers take as input whitespace-delimited tokens. However, in Arabic word boundaries can be more complex.
Arabic letters can have up to 
four different forms
corresponding to their position in a word, i.e. beginning,
middle, end of word and separate forms. 
This allows the phrase
\noTrnoVocRL{il_A\nospace almdrsT}  with no delimiter within
to be visually recognizable
as two separate words \noVocRL{il_A} (to) and \noVocRL{almdrsT} (the school).
This happens because the first word \noVocRL{il_A} ends with
\noVocRL{_A} a non-connecting letter. We propose to solve this problem, commonly known as the run-on word problem, using online tokenization.
This also solves the issue of multi-word expressions such as \noVocRL{`abid al-la.tiyf} being analyzed as one unit -- here a compound proper name.

%Arabic however has distinctive characteristics that
%complicate its analysis. 
%Arabic text is composed of

In this thesis, we propose to build Sarf, an application-specific 
morphological analyzer for the Arabic language 
which is capable of benefiting
from the specific characteristics of the application at hand.
In addition , Sarf solves the issue of `run-on' words, supports disambiguation using diacritics, 
handles multi-word expressions and performs online tokenization. 
We will apply Sarf to  
(1) extract time entities from Arabic documents,
(2) extract narrators from hadith documents,
(3) relate narrators to their biographies in separate documents, and 
 thus automate the process of hadith authentication.

\subsubsection{Time Entity Extraction}

%One of the applications we propose to build using Sarf is a time entity extractor. This application
%recognizes time expressions in text document and transforms them to time information of the default format
%that can be stored in a database and on which computations and analysis can be performed.
Consider the following phrase: \noVocRL{al-'i_tnayn fy al-sAbi` min ^sbA.t lsanaT 2011} (Monday the seventh of February 2011).
We are interested in transforming such phrases into a time structure of the form: 05/02/2011 that can be used in further analysis.
Many applications can benefit from temporal entities extracted from text documents.
For example, %applications that sort the documents chronologically.
Li et al. from 
Microsoft hold a patent for a tool that extracts authorship 
dates from Latin documents~\cite{patent:09}. We will use Sarf to automate time entity extraction for Arabic.

\subsubsection{Hadith Authentication}

In this thesis, we will present and propose to solve an interesting application 
for text mining in the field of Islamic Studies. 
A \noVocRL{.hady_t}
is a narration related to the prophet Mohammad
through a \noVocRL{sanad} or a sequence of narrators. 
The collections of traditions are the second source of
jurisprudence after the \noVocRL{qor'An} for all Islamic schools of thought. 
Figure~\ref{f:exhadith} shows an example \noArRL{.hady_t} in 
Arabic with its transliteration and translation. 
We show proper names in boxes connected
to form complex names of narrators. 
For example, 
\novocalize
\noTrRL{qtybT} is the first name 
of narrator $n_1$, and 
\noTrRL{s`yd} is the name of his father as 
the word \noTrRL{bn} (son of) indicates. 
The sequence of names from $n_1$ to $n_5$ 
constitutes the \noArRL{sanad} 
of the \noArRL{.hady_t}. 
The second part of the \noArRL{.hady_t} is the 
\RL{matn} (content) and constitutes the actual content 
of the tradition.

@@

\section{Other Applications of Text Mining}
\label{sec:Applications}

Research in text mining started in the mid 1980s when Swanson 
realized that slicing and combining seemingly unrelated medical 
articles led to the discovery of new 
hypotheses~\cite{JNi06}. 
Text Mining has gained much interest lately due to its wide range of applications,
including its use in the areas of advertisement, 
political campaigning, and other businesses.
These applications consider corpora that substantially differ in terminology, 
sentence structure and morphology. Thus, an application-specific analyzer such as Sarf is of great benefit.

In this section, we present previously explored applications of text mining. %TODO: and focus on their discrimenent morphological features.

\subsection{Sentiment Analysis} 
	This field -- also known as `Opinion Mining' -- attempts classifying the polarity of sentiments of people 
	or even their emotional state ~\cite{Pang:08}. The inputs for such analyzers include product and movie reviews, 
	Internet blogs, electronic social networks and possibly newspapers. Early research was conducted on review sites
	to automate the task classifying whether the review was positive and negative. Consequently, users can get concise helpful
	sentiment summaries about the product, movie, etc...~\cite{Pang:02, Turney:02}. Such abilities motivated research targeting 
	the following applications:
	\subsubsection{Marketing Applications} 
		The strength of blogs in disseminating opinions is a two-edged sword that can 
		``make or break a company's reputation''~\cite{hoffman:08a}. Consequently, analyzing customers' opinions as portrayed 
		on online discussion forums should be detected as soon as possible. This input is vital in the company's policy for its
		reputation management. It can accordingly take actions to retain current customers, attract new ones and win back former 
		customers~\cite{hoffman:08b}. Sentiment Mining has also been used to predict customer churns~\cite{marketing:07,marketing:08}% maybe explain churn
		to take appropriate actions. It can help also in producing more effective advertisements based on the interests of the market.
	\subsubsection{Political Applications} 
		Political experts and analysts can also benefit, since opinions extracted from textual data 
		can reveal much even without the need of performing any explicit surveys. According to ~\cite{Gryc:10}, it ``opens up many avenues
		for more accurate and naturalistic large-scale political analysis.'' Some research has been performed on classifying and predicting
		political inclinations~\cite{Durant:07, Yu:08}. This field can help political leaders to better understand the views of their people 
		and possibly perform better campaigns.

\subsection{Academic Applications} 
	In scientific disciplines where written papers and publications contain technical information, the need to retrieve
	specific information that a user needs, aids in \textbf{\em developing advanced search engines}. 
	Text mining plays a great role toward this end through the following:
	\subsubsection{Text Summarization} 
		When an application is interested in mining a text for the most important and informative portions,
		the application is considered to be tackling the field of `Text Summarization'. Summaries may have different forms such as being 
		a group of representative words, or even a coherent abstraction of the text in novel terms and phrases that capture its meaning. 
		The main motivation
		for such a research is to enhance search engines. By capturing the most informative concepts in a document, it aids determining a 
		better metric for a document's relevance to a search query. It also helps search engines display a shortened preview of the document for returned
		results. This reduces the overwhelming amount of information a user needs to read before deciding whether the whole document deserves
		comprehensive study. It is also helpful when the editorial staff of a magazine or newspaper need to present a large article in the limited space.
		Similarly, summarization is vital when the content of 
		news articles needs to be dramatically shortened to be sent in an SMS form to subscribed PDA users. 
		Useful surveys of current research on text Summarization can be
		found in ~\cite{Radev:02,Das:07}.
	\subsubsection{Text Categorization and Clustering} 
		In many cases, it is very useful to be able to search for documents by theme, topic, domain, etc...
		This ability is achieved if documents can be automatically clustered into different categories over which a user can search. It has been also 
		proposed to cluster documents based on biographical data~\cite{Kostoff:99}. Technically speaking the difference between categorization and
		clustering is that the former organizes documents into predefined categories where in the latter the categories arise from the document content
		themselves~\cite{sharp:01}. Many approaches exist in this field as described in~\cite{Berry:03, Aas:99, MUNTEANU:05}.
		\begin{enumerate}
		\item \textbf{Email Classification:} It is noteworthy to mention that categorization can be used to classify emails into different subgroups
			-- such as urgent, work, friends, spam, etc... Consequently, it can be used for spam detection and email 
			filtering~\cite{Mock:01,bekkerman:04}.
		\item \textbf{Parental Control:} It has also been proposed to use categorization for parental control by
			automatically detecting inappropriate content for children~\cite{Kontostathis:09}.
		\end{enumerate}
	\subsubsection{Entity Extraction:} Named Entity Extraction/Recognition (NER) targets being able to detect phrases that convey some general concept
		and/or belong to a predefined set of categories such as persons, places, time, etc... It has been argued that having the luxury of searching
		for entities enhances search abilities instead of being restricted to search for a list of words appearing in a document~\cite{basis:06}. 
		It seems that humans think of their
		queries in terms of entities/concepts and then try translating them to some list of words which are most likely to appear in a document that
		is relevant to their search query. In general, a user does not ``know all the search terms that will return all instances of the desired
		information''~\cite{basis:06}. A survey of research in the field is presented in ~\cite{Nadeau:07}.

\subsection{Security Applications} 
	It has been proposed that text mining of emails can be used for detecting money laundering and 
	terrorist plans. However the effectiveness of mining techniques in this context is still debatable~\cite{Jonas:06,Federici:07}.

	We also propose another security application which can be used when investigating in criminal cases:
	Given two sets of reports consisting of crime scene investigation and suspect interrogation reports, 
	an officer may use a specialized text mining application to query for relations between suspects in terms of 
	criminal action, locality, or third party persons. %common between suspects
	It can be envisioned that such an application will return two convicted people who had no direct relation
	in the reports, but who are related only through a third person unknown to the investigator. 
	Such a query is very hard to answer manually.

\subsection{Medical Applications} 
	Medical records and clinical reports are studied to detect
	patterns of occurrence of certain symptoms and diseases. It attempts to relate those patterns with
	other biological factors such as certain 
	genes or protein interaction~\cite{blaschke:99, Krallinger:05, Rebholz-Schuhmann:07, Cohen:08, Tsuruoka:08}.
	They are also used to analyze factors correlating with medical accidents to minimize their occurrence~\cite{Kimura:08}.
	Till present, most of the interest in text mining comes from biological sciences~\cite{Giannopoulou:08}.

@@

	Morphological ambiguity is a `notorious' problem for the Arabic language~\cite{Kiraz:98}.
	Ambiguity refers to the fact that one word may lead to several possible solutions.
	As discussed in~\cite{Attia:08a}, the following are the reasons of genuine ambiguity\footnote{
	Here we are discussing ambiguity that is inherent to the language. Other sources of ambiguity
	may be due to over-approximations and relaxed compatibility rules in the morphological analyzer.} 
	in the language.
	\begin{enumerate}
		\item \textbf{Orthographic alternation operations:} These refer to alterations from the stem
			original letters upon inflection, which can make it similar to other distinct words.
			For instance, \noVocRL{ya`id} can result in 5 different solutions, 2 of which are
			inflections of \RL{A`id} (bring back) and \RL{`od} (come back) as in \RL{lam ya`od}.
			~\cite{Attia:08a}
		\item \textbf{Prefixes and Suffixes that are homomorphic:} For example, 
			``the prefix {\em ta-} can indicate 3rd person feminine or 2nd person masculine''~\cite{Attia:08a}.
			Consider \noVocRL{tadros} which can mean `she studies' or `you (masculine) study'.
		\item \textbf{``Coincidental Identity'':} In this case, complex forms (i.e. containing affixes) coincidentally
			are homomorphic to a simple full word ~\cite{Kamir:02}. 
			%An example is in the homomorphism of \noVocRL{Akala}
			%he ate) and \noVocRL{Akalla} [\noVocRL{A} + \noVocRL{kalla}] (he made tired).
			\vocalize 
			For instance, the word \RL{'a.hmadH} may have two valid morphological analyses. 
			The word means ``I praise him'' when the letter \RL{'a} is a prefix and  it means
			``his Ahmad'' when the letter \RL{'a} is part of the proper noun stem \RL{'a.hmad}.
		\item \textbf{Natural Homomorphism in Un-inflected Forms:} This occurs in \noVocRL{_dahaba} which can mean
			`went' as a verb and `gold' as a noun.~\cite{Attia:08a}
		\item \textbf{Omitted Diacritizations} (will be discussed next)
	\end{enumerate}

	The statistical measure of ambiguity is the average count of valid solutions per word. According to our results performed
	on 3 books of hadith~\cite{IbnHanbal,AlKulayni,AlTousi}, the resulting ambiguity score was 2.62 using an extended corpus to that of
	Buckwalter Analyzer~\cite{Buckwalter:02}\footnote{This analyzer will be considered in Section \ref{sec:literature}.}.
	Those results where verified by ~\cite{Attia:06b} which reported a similar score (2.6) and a score of 1.76 for English.
	This difference can be used to argue that Arabic is more ambiguous than English. Still however, the author of ~\cite{Attia:06b}, Attia,
	claims that this ambiguity resulted from some grammatical over-approximations and inaccuracies in the lexicon of Buckwalter along with
	its inclusion of classical terms that are rarely used. By this he means that the score does not reflect the intrinsic Arabic ambiguity and 
	he reports a score of 1.75 for his tool which used tighter grammar and lexical rules and omitted classical words. Although, Dr. Attia
	has a point relating to exaggerated numbers due to over-approximations, the omission of some non-frequent words from the lexicon 
	does not a correct approach for reporting ambiguity since it does
	not reflect the ambiguity of Arabic but of the subset of Arabic needed for some applications.\footnote{
	In his case, he focused mainly on modern text documents such as news articles.}
	In all cases, this is a question open to future investigations by NLP researchers and Arabic linguists...

@@

	Since most Arabic documents are unvocalized, the ambiguity of the morphological analyses is significantly increased.
	For instance, in many cases, it is impossible to distinguish between active and passive verbs on the word level without
	appropriate diacritics. For instance the undiacriticized versions of \RL{`abada} ($<$subject$>$ worshiped) and \RL{`obida} ($<$subject$>$ was worshiped) are homomorphic.

@@

%	Arabic letters can have up to 
%	four different forms
%	corresponding to their position in a word, i.e. beginning,
%	middle, end of word and separate forms. 
%	This allows the phrase
	The example provided in the introduction,
	\noTrnoVocRL{il_A\nospace almdrsT} provides evidence to the issues 
	that can arise due to the presence of non-connecting letters. 
	The non-connecting letters are \RL{A}, \noVocRL{'a}, \noVocRL{'i},
		\noVocRL{'A},\noVocRL{_A}, \noVocRL{w}, \noVocRL{|u'|}, \noVocRL{|i'|},	
		\noVocRL{d}, \noVocRL{_d}, \noVocRL{r}, \noVocRL{z} and \noVocRL{T}.
	Words that are not separated by white-space occur often in text and 
	greatly increase the
	difficulty of tokenization and are referred to as 
	``run-on'' words~\cite{Buckwalter:04}.
	Such difficulties are inherent in the 
	morphological analysis of the Arabic language and
	render the solutions inaccurate.
	Consequently, tokenization of Arabic text is not a trivial task of 
	detecting delimiters as is the case in other languages.

@@

The rest of the proposal is organized as follows. In Section 2, we present the problems to be attempted. 
In Section 3, we review related literature. 
Then, in Section 4, we describe our preliminary results. 
Finally, in Section 5, a time-schedule and a research plan are presented.

@@

In other words, if the analyzer can be controlled at every step by an application-specific
controller, the analysis can be enhanced significantly.
